Persistence in functional topology: local-connectivity, q-tameness and a correction to a theorem of Morse

### 2021-04-15 connectivity.tex
Add remark about spherical classes to historical section

Local connectivity assumptions have a long history of being used to ensure that different homology theories agree and to ensure that homology is finite dimensional.
We mention the following examples.

Recall that a space is said to be \textit{paracompact} if every open cover has a locally finite open refinement, i.e., every point has a neighborhood intersecting finitely many sets in the refinement.
\begin{prop}[\cite{MR105677}] \label{prop:cech_sing_hom_hlc}
	\v{C}ech and singular homology with arbitrary coefficients coincide on the category of paracompact Hausdorff spaces which are $\HLC$ with respect to integral singular homology.
\end{prop}

>>>>

Local connectivity assumptions have a long history of being used to ensure that different homology theories agree and to ensure that homology is finite dimensional.
We mention the following examples of these two uses, recalling that a space is said to be \textit{paracompact} if every open cover has a locally finite open refinement, i.e., every point has a neighborhood intersecting finitely many sets in the refinement.

\begin{prop}[\cite{MR105677}] \label{prop:cech_sing_hom_hlc}
	\v{C}ech and singular homology with arbitrary coefficients coincide on the category of paracompact Hausdorff spaces which are $\HLC$ with respect to integral singular homology.
\end{prop}

### 2021-04-15 historical.tex
Add section by section to intro
Add remark about spherical classes to historical section

\section{Historical remarks} \label{sec:historical}

@@

Morse assumes all spaces in his filtrations to be compact and uses Vietoris homology, which agrees with \v{C}ech homology by Dowker's Theorem \cite{Dowker.1952}. This implies semi-continuity by a general result stating that, for field coefficients, \v{C}ech homology commutes with inverse limits for compact Hausdorff spaces \cite[Theorem VIII.3.6 and Theorem X.3.1]{MR0050886}. To achieve q-tameness, Morse wants to use the following condition, which he refers to as \textit{local $F$-connectivity}. 
\begin{displaycquote}[p.431]{Morse.1940}
	The space $M$ is said to be \textit{locally $F$-connected} of order $r$ at $p$ if corresponding to each positive constant $e$ there exists a positive constant $\delta$ such that each singular $r$-sphere on the $\delta$-neighborhood of $p$ on $F_{c+\delta}$ bounds an $(r+1)$-cell of norm $e$ on $F_{c+e}$.
\end{displaycquote}
Morse then states that as consequence of bounded compactness and local $F$-connectivity the rank of the persistence Vietoris homology module associated to $(M, F)$ is q-tame.
In the original it reads:
\begin{displaycquote}[Theorem 6.3, p.432]{Morse.1940}
	Let $a$ and $c$ be positive constants such that $a < c < 1$.
	The $k^{\mathrm{th}}$ connectivity $R^k(a,c)$ of $F_a$ on $F_c$ is finite.
\end{displaycquote}
The proof Morse cites, found in \cite[Theorem 6.1]{Morse.1938}, is not correct; neither is the statement.
A stronger notion of local connectivity for the pair $(M, F)$ is needed to prove q-tameness.
Such condition had already appeared in print, introduced by Morse himself, three years earlier.
\begin{displaycquote}[p.421-422]{Morse.1937}
	The space $M$ will be said to be locally $F$-connected for the order $n$ if corresponding to $n$, an arbitrary point $p$ on $M$, and an arbitrary positive constant $e$, there exists a positive constant $\delta$ with the following property. For $c \geq F(p)$ any singular $n$-sphere on $F \leq c$ (the continuous image on $F \leq c$ of an ordinary $n$-sphere) on the $\delta$-neighborhood $p_{\delta}$ of $p$ is the boundary of a singular $(n + 1)$-cell on $F \leq c + e$ and on $p_e$.
\end{displaycquote}
This notion of local-$F$-connectedness is stronger than the one first quoted, with the key difference being the freedom granted by the constant $c \geq F(p)$.
Following the introduction of this stronger definition Morse states as \cite[Theorem~9.2, p.422]{Morse.1937} that q-tameness follows from this form of local $F$-connectivity and bounded compactness, adding that the proof ``while not difficult involves too great detail to be presented here".

>>>>

\section{Historical remarks} \label{s:historical}

@@

Morse assumes the spaces under consideration are equipped with a metric -- in particular are locally compact -- and uses Vietoris homology, which agrees with \v{C}ech homology by Dowker's Theorem \cite{Dowker.1952}.
This implies semi-continuity by a general result stating that, for field coefficients, \v{C}ech homology commutes with inverse limits for compact Hausdorff spaces \cite[Theorem VIII.3.6 and Theorem X.3.1]{MR0050886}.
To achieve q-tameness, Morse wants to use the following condition, which he refers to as \textit{local $F$-connectivity}. 
\begin{displaycquote}[p.431]{Morse.1940}
	The space $M$ is said to be \textit{locally $F$-connected} of order $r$ at $p$ if corresponding to each positive constant $e$ there exists a positive constant $\delta$ such that each singular $r$-sphere on the $\delta$-neighborhood of $p$ on $F_{c+\delta}$ bounds an $(r+1)$-cell of norm $e$ on $F_{c+e}$.
\end{displaycquote}
Morse then states that as consequence of bounded compactness and local $F$-connectivity the rank of the persistence Vietoris homology module associated to $(M, F)$ is q-tame.
In the original it reads:
\begin{displaycquote}[Theorem 6.3, p.432]{Morse.1940}
	Let $a$ and $c$ be positive constants such that $a < c < 1$.
	The $k^{\mathrm{th}}$ connectivity $R^k(a,c)$ of $F_a$ on $F_c$ is finite.
\end{displaycquote}
The proof Morse cites, found in \cite[Theorem 6.1]{Morse.1938}, is not correct; neither is the statement.
A stronger notion of local connectivity for the pair $(M, F)$ appeared in print, introduced by Morse himself three years earlier.
\begin{displaycquote}[p.421-422]{Morse.1937}
	The space $M$ will be said to be locally $F$-connected for the order $n$ if corresponding to $n$, an arbitrary point $p$ on $M$, and an arbitrary positive constant $e$, there exists a positive constant $\delta$ with the following property. For $c \geq F(p)$ any singular $n$-sphere on $F \leq c$ (the continuous image on $F \leq c$ of an ordinary $n$-sphere) on the $\delta$-neighborhood $p_{\delta}$ of $p$ is the boundary of a singular $(n + 1)$-cell on $F \leq c + e$ and on $p_e$.
\end{displaycquote}
This notion of local-$F$-connectedness is stronger than the one first quoted, with the key difference being the freedom granted by the constant $c \geq F(p)$, just as in our definition of $\HLC$ and weakly $\HLC$ presented in Definition~\ref{defi:local_connectedness_filtrations}.
Following the introduction of this stronger definition Morse states as \cite[Theorem~9.2, p.422]{Morse.1937} that q-tameness follows from this form of local $F$-connectivity and bounded compactness, adding that the proof ``while not difficult involves too great detail to be presented here".
This claim is not true since not all singular homology classes can be represented by the pushing forward the fundamental class of spheres.
We could use a version of the Hawaiian earring made out of tori instead of spheres to construct a counterexample.
